% ************************** Thesis Acknowledgements **************************

\begin{acknowledgements}      

First of all, I would like to express my gratitude to Dr Paul Griffiths and Prof Alexander Archibald for their insights, advice, patience, and guidance. Thank you for being so generous with your time, effort, and ideas at every stage of this research. It has been so rewarding working with you, and I could not have done it without your support.

I would like to thank the Cambridge Thai Foundation and the Cambridge Trust for funding this study. The UKRI Centre for Doctoral Training (CDT) in the Application of Artificial Intelligence to the Study of Environmental Risks (AI4ER) generously accepted me as an aligned student in the programme, from which I benefited from training and peer support.

I would like to thank all CAS members for welcoming me into the group and for both academic and technical advice. Thank you to Dr Luke Abraham for his role as my academic mentor. Our check-ins have been a steadying presence throughout. Special thanks to all my PhD student peers: Zosia Staniaszek, Lorrie Jacob, May Chim, Sophie Turner, Seb Hickman, and Michelle Wan, for their friendship and great company throughout the journey. I may be one of the last of us to reach the finish line, but I will make it there, and you all have shown me it is possible! I am also grateful to Megan Brown and Kita for delightful excursions and moral support. Through the enthusiasm of Charlie Wartnaby and Rob Waters, I have embraced park running and code reviews as part of my life. Thanks to Yao Ge and Laura Stecher for keeping me company through the writing-up and the transition from a nearly-finished PhD to life as a PDRA.

I have transformed throughout the PhD journey, hitting numerous rock bottoms. I would not have bounced back without all the people I am so grateful to have in my life. This part of the thesis may be the one I was most looking forward to writing.

Paul, not only are you my supervisor, but I also think of you as a good friend and mentor. I do not think I would have continued or finished this PhD without your encouragement and support. The many hours you have spent working with me will form the foundation of the type of mentor I would like to be (if I ever become one).

I am grateful for my friends at St Edmund's College, who have become my away-from-home family. Being part of the Student Committee through the first year of the COVID pandemic has been transformative. Chris, thank you for showing so much leadership and warm support. Your editing skills are nothing short of amazing. Thanks to Abigail, Elsa, Zeke, and the no-longer-little Lucy, for all the playtime too! Linea, thanks for all the fun activities, welcoming me into your home, and for exposing me to knitting. Copenhagen is now my favourite capital city because you are there. Madan and Silvia, thank you for keeping me company and for the many outings and fun chats. Devathi, our days together at College may have been short, but thank you for all the chats on Instagram and the postcards; I kept them all. 

To my housemates, I am grateful to have you as my friends. Fynn, you are such an amazing and caring person. Thank you for being there, for listening to all my rambles, and for all the impromptu Kerb trips. James, I am glad we get to share so many house dinners and shows together.

I am grateful to have had access to the University Counselling Service, which has supported my mental well-being. Thank you to Kyoko, who has helped me introspect on my relationship with writing. I would also like to dedicate a special thank-you to my long-time, now former, psychiatrist, Dr Puchong, who has challenged my unhelpful and biased thoughts. The skills I learned have brought me to this point, and I will carry with me what I learned from you.

I am also grateful for the wider community around me. Thank you to Kerb Collective and Maison Clement for offering the best coffee that fuelled my writing and for the sweet treats when I needed a pick-me-up.

To my friends and family here in the UK and back home, especially Pom, Jang, and Praifa, thanks for all the birdwatching and board game days that kept me sane. Yanin, Bow, Bas, Luknam, Gift, Nueng, Jane, Ploy, and all my TU friends --- there are too many of you to name --- for our annual gatherings and for believing that I will eventually finish the PhD. To my NARIT friends, especially Tuck and Toothless, for letting me visit every year. To Klod, for your messages of encouragement. To Jittip, for countless trans-Atlantic co-working videocalls. To Krzysztof, for letting me into your life. To Hannah, my cat, who did nothing to help, but whose existence cheers me up. Finally, to my sister, Pichalai, for her encouragement and unwavering support.

\end{acknowledgements}
